Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-06-22 GWLDCII.tex
fixed a whole in the purple proof

Write $I_i(p_j)=a_j$. By algorithm \ref{alg:put GN on WLD}, $a_j$ is the smallest number in the $<_{a_{j-1}}$ ordering such that $c_{j, a_j} \neq 0$ (that is, is an indeterminant entry). It is the first column in the $j^{th}$ row of $C(W)$ that is both non-zero and is greater than the index contributed by the previous propagator:  \bas \begin{bmatrix}\ddots & & & & & & \\ \ldots  & c_{j-1, a_{j-1}} & * & \ldots &  & & \\  \ldots  &* &0 & \ldots & 0 &c_{j, a_j} & \ldots \\ & & & & & & \ddots \end{bmatrix}  \eas

\hlfix{In particular, any number between two consecutive indices in $I_i$, say $a_j < v < a_{j+1}$, correspond to vertices inside the propagator $p_j$ in the $a_j-1$ ordering. }{I know we got rid or interior vertices. Can we bring it back or rephrase?} Furthermore, every propagator preceeding $p_j$ is outside it in the same ordering: $\cP_{j-1} \subset \cP_{out}(p_j)$ in the $a_j-1$ ordering.

Let $J = (j_1,\ldots j_k)$ be a sequence that is lexicographically less than $I_i = (a_1 \ldots a_k)$: $J <_i I_i$. By the above, for some $m < k$, $j_m <_i a_m$ and the matrix entry $c_{m, j_m} = 0$ is $0$. Without loss of generality, let $m$ be the smallest $m$ in the $<_i$ order for which this is true: \bas \forall l <m, \; I_i(p_l) = j_l \; ; \; j_m < I_i(p_m) \;.\eas Write $J_m = (j_1, \ldots, j_m)$ indicate the first $m$ entries of $m$.

We claim that $J_m$ is not a basis for $W|_{\cP_m}$. To see this, first note that, since $c_{m, j_m} = 0$, the propagator $p_m$ is not supported on $j_m$. Furthermore, by construction, $a_{m-1} < j_m < a_m$. Therefore, $j_m \in V_{in}(p_{j_m})$ in the $a_j-1$ ordering. Since  $\cP_{j-1} \subset \cP_{out}(p_j)$ in the $  a_j-1$ ordering, non-crossing of propagators ensures that none of the first $m-1$ propoagators are supported on the vertex $j_m$. This implies that the minor of $C(W|_{\cP_m})$ defined by $J_{m}$, has a column of $0$s. That is, $J_m$ is not an independent set in $C(W|_{\cP_m})$, and thus not in $C(W)$.

\end{proof}
}
We can now unambiguously refer to Algorithm~\ref{alg:put GN on WLD} as the Grassmann necklace algorithm. We close this section with a simple observation about the Grassmann necklaces of positroids associated to Wilson loop diagrams.

>>>>

Write $I_i(p_j)=a_j$. By algorithm \ref{alg:put GN on WLD}, $a_j$ is the smallest number in the $<_{a_{j-1}}$ ordering such that $c_{j, a_j} \neq 0$ (that is, is an indeterminant entry). It is the first column in the $j^{th}$ row of $C(W)$ that is both non-zero and is greater than the index contributed by the previous propagator:  \bas \begin{bmatrix}\ddots & & & & & & \\ \ldots  & c_{j-1, a_{j-1}} & * & \ldots &  & & \\  \ldots  &* &0 & \ldots & 0 &c_{j, a_j} & \ldots \\ & & & & & & \ddots \end{bmatrix}  \eas
Since $p$ never contributes its last vertex in the $<_i$ order to $I_i$, by Lemma \ref{lem no fourth vertex}, every propagator preceeding $p_j$ is outside it in the $a_j-1$ ordering: $\cP_{j-1} \subset \cP_{out}(p_j)$.

Let $J = (j_1,\ldots j_k)$ be a sequence that is lexicographically less than $I_i = (a_1 \ldots a_k)$: $J <_i I_i$. By the above, for some $m < k$, $j_m <_i a_m$ and the matrix entry $c_{m, j_m} = 0$ is $0$. Without loss of generality, let $m$ be the smallest $m$ in the $<_i$ order for which this is true: \bas \forall l <m, \; I_i(p_l) = j_l \; ; \; j_m < I_i(p_m) \;.\eas Write $J_m = (j_1, \ldots, j_m)$ indicate the first $m$ entries of $J$.

We claim that $J_m$ is not a basis for $W|_{\cP_m}$. To see this, first note that, since $c_{m, j_m} = 0$, the propagator $p_m$ is not supported on $j_m$. If $p_m$ does not share support with any propagator in $\cP_m$ in the cyclic interval $[i, j_m]$, then the minor of $C(W|_{\cP_m})$ defined by $J_{m}$, has a row of $0$s.

Otherwise, let $p_l$ be a propagator preceeding $p_m$ ($p_l \in \cP_{m-1}$) that shares support with it in $J_m$: $V(p_l) \cap V(p_m) \cap J_m \neq \emptyset$. Then $p_l \in Q$ is positioned clockwise to $p_m$ on vertex $a_l$, therefore, $p_l$ is outside $p_m$ in the $a_l -1$ order. Since $a_l <_i j_m <_i a_m$ with $a_l$ and $a_m$ both in the support of $p_m$, \hlfix{$j_m$ is inside $p_m$ in the $a_l-1$ ordering. }{I know we got rid or interior vertices. Can we bring it back or rephrase?} Furthermore since all of $\cP_{m-1} \subset \cP_{out}(p_m)$ is outside of $p_m$ in that $a_l-1$ order, and $j_m$ is a vertex inside it, no propagator prior to $p_m$ can be supported on $j_m$. Otherwise, one would have crossing propagators. Therefore, the minor of $C(W|_{\cP_m})$ defined by $J_{m}$, has a column of $0$s.

\end{proof}
}

We can now unambiguously refer to Algorithm~\ref{alg:put GN on WLD} as the Grassmann necklace algorithm. We close this section with a simple observation about the Grassmann necklaces of positroids associated to Wilson loop diagrams.
